\Large{Adversarial Training against Systematic Uncertainty}

### 2016-10-24 tex/paper.tex
save

In this setting, the nuisance parameter $Z$ is considered as a random variable
of prior $p(z)$ (for $z \in {\cal Z}$), and our goal is to find a function
$f(\cdot;\theta_f)$ such that $f(X;\theta_f)$ and $Z$ are independent random
variables.   Importantly, classification of $Y$ with respect to $X$ is
considered in the context where $Z$ is marginalized out, which means that the
classifier minimizing ${\cal L}_f$ is optimal with respect to $Y|X$, but not
necessarily with $Y|X,Z$. Results hold for a nuisance parameter $Z$ taking
either categorical or continuous values. By
abuse of notation, $H(Z)$ denotes the differential entropy in this latter case.
Finally, the  proposition below is derived in a non-parametric setting, by
assuming that both $f$ and $r$ have enough capacity.

@@

In collisions at the Large Hadron Collider, massive particles can be produced at
such high energies that their decay products are collimated and the resulting
jets overlap. Still, deducing whether an observed jet is either due  to a single
hadronic particle or to the decay of a massive object into multiple hadronic
particles  is critical for understanding the nature of the particles produced in
the collision \citep{butterworth2008jet,guest2016jet}. More importantly, this
classification problem is often made difficult by the presence of simultaneous
interactions, known as pileup, which contributes significant energy depositions
unrelated to the particles of interest.

In this setting, we reuse the classification problem of \citep{guest2016jet}
between single jets produced in quark or gluon fragmentation and two overlapping
jets produced when a high-velocity $W$ boson decays to a collimated pair of
quarks. As nuisance parameter, we consider events without pileup ($Z=0$) and
events with pileup ($Z=1$), for which an average number of $\langle \mu \rangle =
50$ unrelated interactions are overlaid. Our goal is to build an accurate
classifier, for which we also want to minimize the effects due to the
uncertainties on the nuisance. More specifically, we choose to recast the
classification problem as a hypothesis test between signal+background (jets
originating from $W$ bosons or from single quarks and gluons) and background
only (jets originating from single quarks and gluons only) and tune the decision
threshold of our classifier by maximizing its approximate median significance
(AMS), when uncertainties in the background are taken into account (see Eqn.~20
of \citep{adam2014higgs}). Our motivation is that reducing the effects of
uncertainties by requiring independence of $Z$ with the classifier output should
allow for a larger maximum significance.

@@

As shown in  Fig.~\ref{fig:physics-ams}, trading classification accuracy for
independence with respect to pileup results in a positive total net benefit in
terms of statistical significance. Without adversarial training (at
$\lambda=0$), the maximum significance peaks at $7$. By contrast, as the
independence constraint is made stronger (for $\lambda > 0$) the AMS peak moves
higher, with a maximum value around  $7.8$ for $\lambda=10$. In other words,
sacrificing accuracy to reduce  the  effects of the systematics in the
background jets does indeed allow one to improve the overall significance of the
results. Setting $\lambda$ too high however (e.g. $\lambda=500$) results in a
decrease of the maximum significance, by focusing $f$ too much on independence,
at the expense of too large a drop in accuracy.

>>>>

In this setting, the nuisance parameter $Z$ is considered as a random variable
of prior $p(z)$ (for $z \in {\cal Z}$), and our goal is to find a function
$f(\cdot;\theta_f)$ such that $f(X;\theta_f)$ and $Z$ are independent random
variables.   Importantly, classification of $Y$ with respect to $X$ is
considered in the context where $Z$ is marginalized out, which means that the
classifier minimizing ${\cal L}_f$ is optimal with respect to $Y|X$, but not
necessarily with $Y|X,Z$ (unless $Z$ is made explicit and is included among the
input variables in $X$). Results hold for a nuisance parameter $Z$ taking either
categorical or continuous values. By abuse of notation, $H(Z)$ denotes the
differential entropy in this latter case. Finally, the  proposition below is
derived in a non-parametric setting, by assuming that both $f$ and $r$ have
enough capacity.

@@

In collisions at the Large Hadron Collider, massive particles can be produced at
such high energies that their decay products are collimated and the resulting
jets overlap. Still, deducing whether an observed jet is either due  to a single
hadronic particle or to the decay of a massive object into multiple hadronic
particles  is critical for understanding the nature of the particles produced in
the collision \citep{butterworth2008jet,guest2016jet}. Critically, this
classification problem is often made difficult by the presence of simultaneous
interactions, known as pileup, which contributes significant energy depositions
unrelated to the particles of interest.

In this setting, we reuse the classification problem of \citep{guest2016jet}
between single jets produced in quark or gluon fragmentation and two overlapping
jets produced when a high-velocity $W$ boson decays to a collimated pair of
quarks. As nuisance parameter, we consider events without pileup ($Z=0$) and
events with pileup ($Z=1$), for which an average number of $\langle \mu \rangle =
50$ unrelated interactions are overlaid. Our goal is to build an accurate
classifier, for which we also want to minimize the effects due to the
uncertainties on the nuisance. More specifically, we choose to recast the
classification problem as a hypothesis test between signal+background (jets
originating from $W$ bosons or from single quarks and gluons) and background
only (jets originating from single quarks and gluons only) and tune the decision
threshold of our classifier by maximizing its approximate median significance
(AMS), when uncertainties in the background are taken into account (see Eqn.~20
of \citep{adam2014higgs}). Our motivation is that reducing the effects of
uncertainties by requiring independence of $Z$ with the classifier output $f(X;\theta_f)$ should
allow for a larger maximum significance.

@@

As  Fig.~\ref{fig:physics-ams} illustrates, without adversarial training (at
$\lambda=0$), the maximum significance peaks at $7$. By contrast, as the
independence constraint is made stronger (for $\lambda > 0$) the AMS peak moves
higher, with a maximum value around  $7.8$ for $\lambda=10$. In other words,
trading classification accuracy for independence with respect to pileup results
in a positive total net benefit in terms of statistical significance. Setting
$\lambda$ too high however (e.g. $\lambda=500$) results in a decrease of the
maximum significance, by focusing the capacity of $f$ too strongly on
independence, at the expense of classification accuracy. As demonstrated in this
example, controlling the classification versus pivot trade-off through $\lambda$
therefore gives us a principled and effective approach for maximizing
significance by desensitizing the classifier output $f(X;\theta_f)$ in the most
beneficial way.
